---
title: "MB1 CDI Follow-up Data Reading and Merge"
author: "The ManyBabies Analysis Team"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
---

# Intro

This Rmd is the first preprocessing file for the primary ManyBabies 1 (IDS Preference) dataset. The goal is to get everything into a single datafile that can be used for subsequent analyses.

Data import functions are factored into a helper functions file.

```{r}
source("helper/preprocessing.R")
```

# CDI Data Import

We first import the newly collected CDI data. Original file names did not always convey encoding type (e.g. `.txt` for tab separated data), hence those were first manually renamed to facilitate automated data reading.

The only other issues were:
- a misnamed column in `infantlanglab-utk.ts`, `CDI_agerange` instead of `CDI-agerange`, corrected by hand;
- extra non-needed columns in a few files, ignored by specifying by hand the columns and column types for import (see helper function);
- standardized scores incorrectly reported as numerical ranges instead of numbers, treated as NAs

```{r}
files.cdi <- list.files("data/cdi/")
data.cdi <- lapply(files.cdi, read_cdi) %>%
  bind_rows()
```

# IDS Preference Data Import

We then import the (pre-processed) data from the original Many Babies study on IDS preference.

```{r}
data.ids <- read_csv("data/ids_preference_trial.csv")
```

# Merge

Finally we merge the IDS preference data into the CDI data, so that only IDS preference data from infants whose CDI score was recorded is kept in the final dataset.

```{r}
data.merged <- left_join(data.cdi, data.ids)
```
